### 2020-11-16 misc/solutions/papa-sols.tex
Fixed numbering in Papa solutions

\maketitlenodate{Answers to Selected Exercises in {\bigboldsl Real and Complex Analysis}\footnote{$^*$}{\frenchspacing \eightpt Walter Rudin. 1987. {\eightsl Real and complex analysis, 3rd ed}. McGraw-Hill, Inc., USA.}}{Solutions by}{Marcel K.\ Goh}

@@

\proclaim 1. Does there exist an infinite $\sigma$-algebra which has only countably many members?

The answer is no.

@@

\proclaim 3. Prove that if $f$ is a real function on a measurable space $X$ such that $\{x:f(x)\geq r\}$ is measurable for every rational $r$, then $f$ is measurable.

\proclaim 5. Prove the following statements:
\medskip
\item{a)} If $f:X\rightarrow[-\infty,\infty]$ and $g:x\rightarrow[-\infty,\infty]$ are measurable, then the sets
$$\{x : f(x) < g(x)\}\ \hbox{and}\ \{x : f(x) = g(x)\}$$
are measurable.
\item{b)} The set of points at which a sequence of measurable real-valued functions converges (to a finite limit) is measurable.

\proof Observe that $f(x) < g(x)$ if and only if there exists a rational number $r$ such that $f(x) < r < g(x)$. So
$$\{x : f(x) < g(x)\} = \bigcup_{r\in \Q} \big(\{x : f(x) < r\} \cap \{x : g(x) > r\}\big) = \bigcup_{r\in \Q} \Big(f^{-1}\big([-\infty,r)\big) \cap g^{-1}\big((r,\infty]\big)\Big)$$
is a countable union of measurable sets (since $f$ and $g$ are measurable functions). Then
$$\{x : f(x) = g(x)\} = \{x : f(x) < g(x)\}^c \cap \{x : g(x) < f(x)\}^c$$
is measurable as well, proving (a).

Let $(f_n)$ be a sequence of real-valued measurable functions. Then $(f_n)$ converges at a point $x$ if and only if it is Cauchy at $x$, i.e.\ for any $\epsilon > 0$ there exists $N\in \N$ such that for all $m,n\geq N$, $|f_m(x) - f_n(x)| < \epsilon$. By the Archimedean property, we can replace $\epsilon$ with $1/k$ for some $k\in \N$. So the set of all points at which $(f_n)$ converges can be written thus:
$$\bigcap_{k=1}^\infty \bigcup_{N=1}^\infty \bigcap_{m=1}^\infty \bigcap_{n=1}^\infty \big\{x : |f_m(x) - f_n(x)| < 1/k\big\}$$
Set $g_{m,n} = f_m - f_n$; this is a measurable function. Hence the set
$$\big\{x : |f_m(x) - f_n(x)| < 1/k\big\} = \big\{x : |g_{m,n}(x)| < 1/k\big\} = {g_{m,n}}^{-1}\big((-1/k, 1/k)\big)$$
is measurable for every $m$, $n$, and $k$. We have thus proved that the set of all points at which $(f_n)$ converges is a countable union of measurable sets.\slug

First we prove that ${\cal M}$ is a $\sigma$-algebra and that $\mu$ is a measure.

Now we claim that the measurable functions are those that are constant at all but countably many points.

Lastly, we show that if $f$ is measurable and takes on the value $k$ at all but countably many points, then $\int_E f\d\mu = k$ for all uncountable $E\in {\cal M}$. (If $E$ is countable then $\int_E f\d\mu = 0$.)

\proclaim 7. Suppose $f_n : X\rightarrow[0,\infty]$ is measurable for $n=1,2,3,\ldots$, $f_1\geq f_2\geq f_3\geq\cdots\geq 0$, $f_n(x)\rightarrow f(x)$ as $n\rightarrow\infty$ for every $x\in X$, and $f_1\in L^1(\mu)$. Prove that then
$$\lim_{n\rightarrow\infty} \int_X f_n \d\mu = \int_X f\d\mu$$
and show that this conclusion does {\it not} follow if the condition ``$f_1\in L^1(\mu)$'' is omitted.

\proclaim 8. Put $f_n = \chi_E$ if $n$ is odd, $f_n = 1 - \chi_E$ if $n$ is even. What is the relevance of this example to Fatou's lemma?

>>>>

\maketitlenodate{Answers to Selected Exercises in {\bigboldsl Real and Complex Analysis}\footnote{$^*$}{\frenchspacing \eightpt Walter Rudin. 1966. {\eightsl Real and complex analysis}. McGraw-Hill, Inc., USA.}}{Solutions by}{Marcel K.\ Goh}

@@

\proclaim 2. Put $f_n = \chi_E$ if $n$ is odd, $f_n = 1 - \chi_E$ if $n$ is even. What is the relevance of this example to Fatou's lemma?

\proclaim 3. Suppose $f_n : X\rightarrow[0,\infty]$ is measurable for $n=1,2,3,\ldots$, $f_1\geq f_2\geq f_3\geq\cdots\geq 0$, $f_n(x)\rightarrow f(x)$ as $n\rightarrow\infty$ for every $x\in X$, and $f_1\in L^1(\mu)$. Prove that then
$$\lim_{n\rightarrow\infty} \int_X f_n \d\mu = \int_X f\d\mu$$
and show that this conclusion does {\it not} follow if the condition ``$f_1\in L^1(\mu)$'' is omitted.

\proclaim 4. Prove that if $f$ is a real function on a measurable space $X$ such that $\{x:f(x)\geq r\}$ is measurable for every rational $r$, then $f$ is measurable.

\proclaim 5. Prove that the set of points at which a sequence of measurable real-valued functions converges (to a finite limit) is measurable.

\proof Let $(f_n)$ be a sequence of real-valued measurable functions. Then $(f_n)$ converges at a point $x$ if and only if it is Cauchy at $x$, i.e.\ for any $\epsilon > 0$ there exists $N\in \N$ such that for all $m,n\geq N$, $|f_m(x) - f_n(x)| < \epsilon$. By the Archimedean property, we can replace $\epsilon$ with $1/k$ for some $k\in \N$. So the set of all points at which $(f_n)$ converges can be written thus:
$$\bigcap_{k=1}^\infty \bigcup_{N=1}^\infty \bigcap_{m=1}^\infty \bigcap_{n=1}^\infty \big\{x : |f_m(x) - f_n(x)| < 1/k\big\}$$
Set $g_{m,n} = f_m - f_n$; this is a measurable function. Hence the set
$$\big\{x : |f_m(x) - f_n(x)| < 1/k\big\} = \big\{x : |g_{m,n}(x)| < 1/k\big\} = {g_{m,n}}^{-1}\big((-1/k, 1/k)\big)$$
is measurable for every $m$, $n$, and $k$. We have thus proved that the set of all points at which $(f_n)$ converges is a countable union of measurable sets.\slug

\noindent First we prove that ${\cal M}$ is a $\sigma$-algebra and that $\mu$ is a measure.

\noindent Now we claim that the measurable functions are those that are constant at all but countably many points.

\noindent Lastly, we show that if $f$ is measurable and takes on the value $k$ at all but countably many points, then $\int_E f\d\mu = k$ for all uncountable $E\in {\cal M}$. (If $E$ is countable then $\int_E f\d\mu = 0$.)

\proclaim 7. Does there exist an infinite $\sigma$-algebra which has only countably many members?

\noindent The answer is no.

@@
